---
title: "Homework 2"
author: "Charlotte Sielewicz"
date: '2022-04-06'
output: html_document
---

# Homework 2

```{r include = FALSE}
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
tidymodels_prefer()
```

## Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no age variable in the data set. Add age to the data set.

Assess and describe the distribution of age.

```{r}
age <- abalone$rings + 1.5
age
Abalone <- cbind(abalone, age)
Abalone  
  head()
```

### Distribution ???

## Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r}
set.seed(123)

abalone_split <- initial_split(Abalone, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```

## Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,

    -   `longest_shell` and `diameter`,

    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = abalone_train) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):shucked_weight) %>% 
  step_interact(~ longest_shell:diameter) %>% 
  step_interact(~ shucked_weight:shell_weight)

```

### **Question 4**

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### **Question 5**

Now:

1.  set up an empty workflow,

2.  add the model you created in Question 4, and

3.  add the recipe that you created in Question 3.

Abalone_train \<- as_tibble(abalone_train)

Abalone_train

```{r}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

### **Question 6**

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_train)
```

```{r}
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy()
New <- data.frame(type = "F", longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)

abalone_train_res <- predict(lm_fit, new_data = New)
abalone_train_res
```

## **Question 7**

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).

    ```{r}
    rmse(abalone_train_res, truth = age, estimate = .pred)
    ```

2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).

    ```{r}
    abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
    abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
    abalone_train_res %>% 
      head()
    ```

3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}
abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, 
                estimate = .pred)
```
