---
title: "Homework 2 [PSTAT 131]"
author: "Charlotte Sielewiz"
date: "4/10/2022"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
library(yardstick)
library(dplyr)
tidymodels_prefer()
abalone <-  read.csv("/Users/charlottesielewicz/Documents/GitHub/PSTAT-131/abalone.csv")
```

## Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no age variable in the data set. Add age to the data set.

Assess and describe the distribution of age.

```{r}
age <- abalone$rings + 1.5

Abalone <- cbind(abalone, age)
Abalone  %>%
  head()
```

```{r}
Abalone %>% 
  ggplot(aes(x = age)) +
  geom_histogram(bins = 60) +
  theme_bw()
```

The data shows that most of the abalone in the `abalone` dataset are around 10 years of age.

## Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r}
set.seed(3435) # set a sead to maintain a consistant answer

abalone_split <- initial_split(Abalone, prop = 0.80,
                                strata = age) # Split the data with an 80:20 ratio
abalone_train <- training(abalone_split) # training set
abalone_test <- testing(abalone_split) # testing set
```

## Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,

    -   `longest_shell` and `diameter`,

    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = abalone_train) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):shucked_weight) %>%  #Interactions type~shucked_weight
  step_interact(~ longest_shell:diameter) %>% # Interaction: longest_shell~diameter
  step_interact(~ shucked_weight:shell_weight) %>% # Interactions: shucked_weight~shell_weight
  step_center(all_nominal_predictors()) %>% # Center all predictors
  step_scale(all_nominal_predictors())      # Scale all predictofs
  
```

## Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

## Question 5

Now:

1.  set up an empty workflow,

2.  add the model you created in Question 4, and

3.  add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>% # Empty workflow
  add_model(lm_model) %>%  # Add model from Q4
  add_recipe(abalone_recipe) # Add recipe from Q3
```

## Question 6

Use your fit() object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_train)
```

```{r}
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy()
New <- data.frame(type = "F", longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)
# Set new data

abalone_train_res <- predict(lm_fit, new_data = New) # Use new data to predict outcome
abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
```

## Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).

    ```{r}
    abalone_metrics <- metric_set(rmse, rsq, mae)
    ```

2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).

    ```{r}
    abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
    abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
    abalone_train_res %>%
      head()
    ```

3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

    ```{r}
    abalone_metrics(abalone_train_res, truth = age, 
                    estimate = .pred)
    ```

The $R^2$ value here is 0.5531....
